Standardized crypto-loans on\\ the Cardano blockchain

### 2021-02-04 papers/actus2021/article.tex
layout (v. minor)

\noindent 
\begin{equation}
\Delta_{r}=\ident{capfloor}(\ident{driver}*\ident{multiplier}+\ident{spread}-\ident{interestRate}_{t-1})
\end{equation}

\noindent 
\begin{equation}
\ident{interestRate}_{t}=\ident{capfloor}(\ident{interestRate}_{t-1}+\Delta_{r}),
\end{equation}

@@

\noindent 
\begin{equation}
\ident{capfloor}(x)=\ident{max(min}(x,\ident{floor}),\ident{cap})
\end{equation}

@@

\noindent 
\begin{equation}
interestPayment=interestScalingFactor*interestRate*notional
\end{equation}
\noindent
This scaling is dynamic and loosely adjusts for variance (volatility)
of the asset that the interest rate driver represents.

@@

\noindent 
\begin{equation}
\ident{payoff}_{i}=\ident{POF}(\ident{state}_{i})
\end{equation}

\newcommand{\STF}{\ident{STF}}
\noindent 
\begin{equation}
\ident{path}=\STF_{1}(ct,ev_{1})\circ \STF_{2}(ct,ev_{2}).\circ\ldots\circ \STF_{i}(ct,ev_{n})
\end{equation}

\noindent 
\begin{equation}
\ident{state}_{i}=\ident{path}(\ident{INIT}(ct)),
\end{equation}
\noindent
where \emph{ct} stands for contract terms, \emph{INIT} returns initial state, \emph{sched}
returns scheduled events, \emph{STF} takes \emph{contractterms}, \emph{event}, and \emph{state}
and returns the next state, and \emph{POF} returns the payoff in a state.

@@

% SJT - the "i" and "t" subscripts here look a bit odd: is that what's intended?
% dk14 - fixed
\noindent 
\begin{equation}
\ident{riskfactor}_{it}=O_{\footnotesize\ident{rf}}(i,t)
\end{equation}
\noindent
In the case of the Cardano blockchain, these values are usually
provided through an \emph{oracle} mechanism\cite{oracles}. An oracle
could be a trusted party providing necessary data or network of parties
under consensus \cite{de-oracles}. 
% SJT REF needed
% dk14. I've found this one has explanation for oracles (ChainLink etc)

@@

\begin{equation}
\ident{formula}(st)=\begin{cases}
\ident{formula}^\prime(st) & \ident{var}_1(st)<\ident{var}_2(st)\\
0 & \ident{otherwise}
\end{cases}
\end{equation}
We rely on a pseudo-Boolean \emph{less than} function in order to address
that:

\noindent 
\begin{equation}
\ident{formula}(st)=\ident{pseudoLt}(\ident{var}_1(st),\ident{var}_2(st)))*\ident{formula}^\prime
\end{equation}

\noindent 
\begin{equation}
\ident{pseudoLt}(a,b)=\ident{Cond}(a>b,1,0)
\end{equation}

@@

% SJT - do you mean \mapsto or \rightarrow here?
% dk14 - it is a mapping, but could be function/morphism too
\noindent 
\begin{equation}
\ident{Applicability}:\ident{ContractTerms}\rightarrow \ident{Bool}
\end{equation}

@@

\smallskip
% SJT changed ContractTerm to ContractTerms
% dk14 OK
\noindent 
\begin{equation}
\ident{Applicability} : \ident{ContractType} \times \ident{ContractTerms} \rightarrow \ident{ApplicabilityType}
\end{equation}
\noindent
where applicability could be: \emph{none}, \emph{always}, \emph{nullable}, or \emph{multiple}.

@@

\noindent 
\begin{equation}
\ident{weaken}(a_{1},a_{2})=\begin{cases}
\ident{nullable} & a_{1}=\ident{none}\wedge a_{2}=\ident{always}\\
\ident{nullable} & a_{1}=\ident{always}\wedge a_{2}=\ident{none}\\
a_{1} & \ident{priority}(a_{1})>\ident{priority}(a_{2})\\
a_{2} & \ident{otherwise}
\end{cases}
\end{equation}

\noindent 
\begin{equation}
\ident{priority}(x)=\begin{cases}
0 & x=\ident{always}\\
1 & x=\ident{none}\\
2 & x=\ident{nullable}\\
3 & x=\ident{multiple}
\end{cases}
\end{equation}
\noindent
where $a_{i}$ is the applicability of a given term of the \emph{i}th contract
type to be merged.

@@

\noindent 
\begin{equation}
\ident{chainlink}(t)=\ident{receiveData}(t)\circ \ident{calculatePayoff(t)}\circ \ident{processPayoff(t)}
\end{equation}

\noindent 
\begin{equation}
\ident{contract}(ct)=\ident{collaterals}(ct)\circ \ident{INIT}(ct)\circ\prod_{\footnotesize t\in \ident{SCHED(ct)}}\ident{chainlink(t)}
\end{equation}

@@

\noindent 
\begin{equation}
\ident{adjusted}=\ident{max(min(original,floor),cap)}
\end{equation}
\noindent
We addressed this issue by introducing the \texttt{Cond} expression construct in order
to represent conditional expressions, rather than only conditional contracts as was the case before. Instead of using the \texttt{If}
contract to decide the value of some variable, we use a conditional expression instead.  
\texttt{Cond} is  a pure function
that returns a value depending on a condition, in contrast to the \texttt{If}
contract that chooses between two continuation contracts. 
% SJT I don't think we need to say this.
%The name is borrowed from a similar function in Lisp.
% dk14 - Oh sure. I was going to remove it too :)

@@

% In a long-term perspective
There is a possible workaround: contracts with no maturity date
could be represented as actors with a finite number of state transitions.
% SJT is this ^^^ rephrasing OK?
% dk14 - sure
We prototyped this approach, % in a separate branch, 
however it does
seem more prone to errors comparing to rendering predefined schedules.
More importantly, it greatly affects static analysis because the number
of reduction steps in the contract grows from 
\noindent 
\begin{equation*}
N_{\footnotesize \ident{scheduled}}=\ident{count(event)}
\end{equation*}
to 
\noindent 
\begin{equation}
N_{\footnotesize \ident{stateTransitions}}=\frac{\ident{max(date(event))}-\ident{min(date(event))}}{\ident{precision}} 
\end{equation}

@@

The Marlowe DSL does not support any notion of records,
the variables can only be of type ``Integer''. In order to
map contract state -- \emph{ContractStatePoly} -- we pack a set of Marlowe
variables of type \texttt{Value} and \texttt{Observation}, representing the previous ($t-1$) state, passing them into
\emph{ContractStatePoly}, to apply the polymorphic state transition, and
finally unpack \emph{ContractStatePoly} into a set of Marlowe variables representing the 
 state at $t+1$:
% SJT again I am confused about t-1 to t+1 transition. Why not just one step?  
% dk14. I'll change that. I imagined t as time of event while t+1 is after event 
\noindent 
\begin{equation}
st_{t}=unpack(stateTransition(pack(st_{t-1}))),
\end{equation}
\noindent
where \emph{pack} is a chain of \texttt{UseValue} constructs and \ident{unpack} is a chain of
\texttt{Let}s.

@@

\newcommand{\rf}{\ident{rf}}
\newcommand{\ct}{\ident{ct}}
\smallskip
\noindent 
\begin{equation}
\forall \ct.\forall \rf. \; \ident{getCashFlows}(\ident{"haskell"},\ct,\rf)\equiv \ident{getCashFlows}(\ident{"java"},\ct,\rf)
\end{equation}
\noindent
where \emph{ct} represents contract terms, \emph{rf} is risk factor model, and
\emph{getCashFlows()} returns a set of \emph{(date,payoff)} tuples.

>>>>

\noindent 
\begin{equation*}
\Delta_{r}=\ident{capfloor}(\ident{driver}*\ident{multiplier}+\ident{spread}-\ident{interestRate}_{t-1})
\end{equation*}

\noindent 
\begin{equation*}
\ident{interestRate}_{t}=\ident{capfloor}(\ident{interestRate}_{t-1}+\Delta_{r}),
\end{equation*}

@@

\noindent 
\begin{equation*}
\ident{capfloor}(x)=\ident{max(min}(x,\ident{floor}),\ident{cap})
\end{equation*}

@@

\noindent 
\begin{equation*}
interestPayment=interestScalingFactor*interestRate*notional
\end{equation*}
\noindent
This scaling is dynamic and loosely adjusts for variance (volatility)
of the asset that the interest rate driver represents.

@@

\noindent 
\begin{equation*}
\ident{payoff}_{i}=\ident{POF}(\ident{state}_{i})
\end{equation*}

\newcommand{\STF}{\ident{STF}}
\noindent 
\begin{equation*}
\ident{path}=\STF_{1}(ct,ev_{1})\circ \STF_{2}(ct,ev_{2}).\circ\ldots\circ \STF_{i}(ct,ev_{n})
\end{equation*}

\noindent 
\begin{equation*}
\ident{state}_{i}=\ident{path}(\ident{INIT}(ct)),
\end{equation*}
\noindent
where \emph{ct} stands for contract terms, \emph{INIT} returns initial state, \emph{sched}
returns scheduled events, \emph{STF} takes \emph{contractterms}, \emph{event}, and \emph{state}
and returns the next state, and \emph{POF} returns the payoff in a state.

@@

% SJT - the "i" and "t" subscripts here look a bit odd: is that what's intended?
% dk14 - fixed
\noindent 
\begin{equation*}
\ident{riskfactor}_{it}=O_{\footnotesize\ident{rf}}(i,t)
\end{equation*}
\noindent
In the case of the Cardano blockchain, these values are usually
provided through an \emph{oracle} mechanism\cite{oracles}. An oracle
could be a trusted party providing necessary data or network of parties
under consensus \cite{de-oracles}. 
% SJT REF needed
% dk14. I've found this one has explanation for oracles (ChainLink etc)

@@

\begin{equation*}
\ident{formula}(st)=\begin{cases}
\ident{formula}^\prime(st) & \ident{var}_1(st)<\ident{var}_2(st)\\
0 & \ident{otherwise}
\end{cases}
\end{equation*}
We rely on a pseudo-Boolean \emph{less than} function in order to address
that:

\noindent 
\begin{equation*}
\ident{formula}(st)=\ident{pseudoLt}(\ident{var}_1(st),\ident{var}_2(st)))*\ident{formula}^\prime
\end{equation*}

\noindent 
\begin{equation*}
\ident{pseudoLt}(a,b)=\ident{Cond}(a>b,1,0)
\end{equation*}

@@

% SJT - do you mean \mapsto or \rightarrow here?
% dk14 - it is a mapping, but could be function/morphism too
\noindent 
\begin{equation*}
\ident{Applicability}:\ident{ContractTerms}\rightarrow \ident{Bool}
\end{equation*}

@@

\smallskip
% SJT changed ContractTerm to ContractTerms
% dk14 OK
\noindent 
\begin{equation*}
\ident{Applicability} : \ident{ContractType} \times \ident{ContractTerms} \rightarrow \ident{ApplicabilityType}
\end{equation*}
\noindent
where applicability could be: \emph{none}, \emph{always}, \emph{nullable}, or \emph{multiple}.

@@

\noindent 
\begin{equation*}
\ident{weaken}(a_{1},a_{2})=\begin{cases}
\ident{nullable} & a_{1}=\ident{none}\wedge a_{2}=\ident{always}\\
\ident{nullable} & a_{1}=\ident{always}\wedge a_{2}=\ident{none}\\
a_{1} & \ident{priority}(a_{1})>\ident{priority}(a_{2})\\
a_{2} & \ident{otherwise}
\end{cases}
\end{equation*}

\noindent 
\begin{equation*}
\ident{priority}(x)=\begin{cases}
0 & x=\ident{always}\\
1 & x=\ident{none}\\
2 & x=\ident{nullable}\\
3 & x=\ident{multiple}
\end{cases}
\end{equation*}
\noindent
where $a_{i}$ is the applicability of a given term of the \emph{i}th contract
type to be merged.

@@

\noindent 
\begin{equation*}
\ident{chainlink}(t)=\ident{receiveData}(t)\circ \ident{calculatePayoff(t)}\circ \ident{processPayoff(t)}
\end{equation*}

\noindent 
\begin{equation*}
\ident{contract}(ct)=\ident{collaterals}(ct)\circ \ident{INIT}(ct)\circ\prod_{\footnotesize t\in \ident{SCHED(ct)}}\ident{chainlink(t)}
\end{equation*}

@@

\noindent 
\begin{equation*}
\ident{adjusted}=\ident{max(min(original,floor),cap)}
\end{equation*}
\noindent
We addressed this issue by introducing the \texttt{Cond} expression construct in order
to represent conditional expressions, rather than only conditional contracts as was the case before. Instead of using the \texttt{If}
contract to decide the value of some variable, we use a conditional expression instead.  
\texttt{Cond} is  a pure function
that returns a value depending on a condition, in contrast to the \texttt{If}
contract that chooses between two continuation contracts. 
% SJT I don't think we need to say this.
%The name is borrowed from a similar function in Lisp.
% dk14 - Oh sure. I was going to remove it too :)

@@

% In a long-term perspective
There is a possible workaround: contracts with no maturity date
could be represented as actors with a finite number of state transitions.
% SJT is this ^^^ rephrasing OK?
% dk14 - sure
We prototyped this approach, % in a separate branch, 
however it does
seem more prone to errors comparing to rendering predefined schedules.
More importantly, it greatly affects static analysis because the number
of reduction steps in the contract grows from 
\noindent 
\begin{equation*}
N_{\footnotesize \ident{scheduled}}=\ident{count(event)}
\end{equation*}
to 
\noindent 
\begin{equation*}
N_{\footnotesize \ident{stateTransitions}}=\frac{\ident{max(date(event))}-\ident{min(date(event))}}{\ident{precision}} 
\end{equation*}

@@

The Marlowe DSL does not support any notion of records,
the variables can only be of type ``Integer''. In order to
map contract state -- \emph{ContractStatePoly} -- we pack a set of Marlowe
variables of type \texttt{Value} and \texttt{Observation}, representing the previous ($t-1$) state, passing them into
\emph{ContractStatePoly}, to apply the polymorphic state transition, and
finally unpack \emph{ContractStatePoly} into a set of Marlowe variables representing the 
 state at $t+1$:
% SJT again I am confused about t-1 to t+1 transition. Why not just one step?  
% dk14. I'll change that. I imagined t as time of event while t+1 is after event 
\noindent 
\begin{equation*}
st_{t}=unpack(stateTransition(pack(st_{t-1}))),
\end{equation*}
\noindent
where \emph{pack} is a chain of \texttt{UseValue} constructs and \ident{unpack} is a chain of
\texttt{Let}s.

@@

\newcommand{\rf}{\ident{rf}}
\newcommand{\ct}{\ident{ct}}
\smallskip
\noindent 
\begin{equation*}
\forall \ct.\forall \rf. \; \ident{getCashFlows}(\ident{"haskell"},\ct,\rf)\equiv \ident{getCashFlows}(\ident{"java"},\ct,\rf)
\end{equation*}
\noindent
where \emph{ct} represents contract terms, \emph{rf} is risk factor model, and
\emph{getCashFlows()} returns a set of \emph{(date,payoff)} tuples.
